Demo arXiv template

### 2022-11-28 paper/paper.qmd
more on the spatial aggregation section

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

```

@@

The construction of an index sometimes needs to consider information from neighbouring time periods.
Temporal processing is a general operator on the time dimension of the data in the form of  $$f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),$$ where $\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}$ is the parameters associated with the temporal operation and $d_{\psi}$ is the number of parameter of $\psi$. 
An typical example of temporal processing is aggregation in the drought index SPI to measure the lack of precipitation for meteorological drought.
In SPI, monthly precipitation is aggregated by a time scale parameter $k$: $x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j),$ where $j^\prime$ is the new time index after the aggregation. In this notation, each spatial location is seaprately aggregated and precipitation is ummed from $k$ month back, $j^\prime - k + 1$, to the current, $j^\prime$ period to create the aggregated series, indexed by $j^\prime$.

@@

**Spatial aggregation** [TODO]

Sometimes

Information from neighbouring spatial locations can also be borrowed when constructing indexes. 
This involves interpolating data from observed locations to the unobserved and aggregating data into different levels.
In the pipeline, spatial aggregation can be written as $$g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),$$ where $\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}$ is the spatial operation parameters and $d_{\theta}$ is the number of parameter of $\theta$. [GIVEN AN EXAMPLE]

While many indexes are calculated independently on spatial locations

satellite images with different resolutions
matching ground and satellite images data

<!-- $x(s_i^\prime;t_j) = \sum_{i \in \Omega}x(s_i; t_j)$ -->

<!-- Another common use of spatial processing is to integrate satellite imagaries with ground measures, also merge satellite imagaries with different resolutions.  -->

 <!-- - 3 by 3 aggregation, another purpose: the spatial grid is too fine, aggregate to higher level. -->

@@

```{r fig-scale-var-trans-compare}
#| fig-cap: "Comparison of scaling and variable transformation."
#| fig-height: 8
#| fig-width: 12
library(tidyverse)
library(latex2exp)
set.seed(123)
x <- rgamma(n = 1000, shape = 1, scale = 2)

@@

Distribution fit can be seen as model fitting in its simplest term. It can be represented by 
<!-- With a probability model imposed,  -->

$$F_{\eta}(x(\mathbf{s}; \mathbf{t})),$$ where $\eta \in H \subseteq \mathbb{R}^{d_{\eta}}$ is the distribution parameter and $d_{\eta}$ is the number of parameter of $\eta$. A distribution fit typically aims at finding the distribution that best fit the data. 
Analysts may start from a pool of candidate distributions with a chosen fitting method and goodness of fit measure. 
While it is useful to find the ultimate best distribution to fits the data, from a probabilistic perspective, the fitting procedure itself has uncertainty associated with the data fed and parameter chosen. A reasonable alternative is to understand how much the index values can vary given different distributions, fitting methods, and goodness of fit tests, and whether these variations are negligible in a given application.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
library(latex2exp)
library(patchwork)
```

@@

The construction of an index sometimes needs to consider information from neighbouring time periods.
Temporal processing is a general operator on the time dimension of the data in the form of  $$f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),$$ where $\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}$ is the parameters associated with the temporal operation and $d_{\psi}$ is the number of parameter of $\psi$. 
An typical example of temporal processing is aggregation in the drought index SPI to measure the lack of precipitation for meteorological drought.
In SPI, monthly precipitation is aggregated by a time scale parameter $k$: $x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j),$ where $j^\prime$ is the new time index after the aggregation. In this notation, each spatial location is seaprately aggregated and precipitation is ummed from $k$ month back, $j^\prime - k + 1$, to the current period, $j^\prime$, to create the aggregated series, indexed by $j^\prime$.

@@

**Spatial aggregation**

While many indexes calculate independently on spatial locations, borrowing spatial information can happen at the unobservable locations. Also, variables can be collected in different formats: some from satellite imageries as raster data, potentially at different resolutions, some from ground stations as vector data. These spatial data need to be merged into a uniform unit to construct indexes. 
In the pipeline, spatial aggregation includes operations on the spatial dimension of the data and can be written as $$g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),$$ where $\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}$ is the spatial operation parameters and $d_{\theta}$ is the number of parameter of $\theta$. 
When one of the variables collected is only available in a coarser resolution, $i^\prime$, analysts may choose to reduce other variables with finer resolution, $i$, to match the coarser resolution, as in @fig-spatial-agg. This is a spatial aggregation step and aggregating using the mean can be written as: $x(s_{i^\prime};t_{j}) = \bar{x}(s_{i \in i^\prime}; t_j),$, where $i \in i^\prime$ includes all the cells from the finer resolution in the coarser grid. *(Feel like the figure here is more of a visual illustration of the equation without enough further implication =.=)*

```{r fig-spatial-agg}
#| fig-cap: "Illustration of aggregating raster data with finer resolution (right) into coarser resolution (left)." 
#| fig-width: 10
#| fig-height: 3
dt <- crossing(y = 1:2, x = 1:3) %>% 
  arrange(-y) %>% 
  mutate(label = as.factor(1:6))
okabe <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#CC79A7")

p1 <- dt %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_tile(aes(fill = label), color = "black") + 
  geom_text(aes(label = label)) + 
  scale_fill_manual(values = okabe) + 
  theme_void() + 
  theme(aspect.ratio = 2/3)

dt2 <- crossing(y = 1:6, x = 1:9) %>% 
  arrange(-y) %>% 
  mutate(x1 = (x-1) %/%3 + 1, y1 = (y-1) %/%3 + 1) %>% 
  left_join(dt, by = c("x1" = "x", "y1" = "y")) %>% 
  mutate(id = row_number())

p2 <- dt2 %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_tile(aes(fill = label), color = "black") + 
  geom_text(aes(label = id)) + 
  scale_fill_manual(values = okabe) + 
  theme_void() + 
  theme(aspect.ratio = 2/3)

(p1 | p2) & theme(legend.position = "none")
```

@@

```{r fig-scale-var-trans-compare}
#| fig-cap: "Comparison of scaling and variable transformation."
#| fig-height: 8
#| fig-width: 12
set.seed(123)
x <- rgamma(n = 1000, shape = 1, scale = 2)

@@

<!-- With a probability model imposed,  -->
Distribution fit can be seen as model fitting in its simplest term. It can be represented by $$F_{\eta}(x(\mathbf{s}; \mathbf{t})),$$ where $\eta \in H \subseteq \mathbb{R}^{d_{\eta}}$ is the distribution parameter and $d_{\eta}$ is the number of parameter of $\eta$. A distribution fit typically aims at finding the distribution that best fit the data. 
Analysts may start from a pool of candidate distributions with a chosen fitting method and goodness of fit measure. 
While it is useful to find the ultimate best distribution to fits the data, from a probabilistic perspective, the fitting procedure itself has uncertainty associated with the data fed and parameter chosen. A reasonable alternative is to understand how much the index values can vary given different distributions, fitting methods, and goodness of fit tests, and whether these variations are negligible in a given application.

### 2022-11-28 paper/paper.tex
more on the spatial aggregation section

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

@@

The construction of an index sometimes needs to consider information
from neighbouring time periods. Temporal processing is a general
operator on the time dimension of the data in the form of
\[f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),\] where
\(\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}\) is the parameters
associated with the temporal operation and \(d_{\psi}\) is the number of
parameter of \(\psi\). An typical example of temporal processing is
aggregation in the drought index SPI to measure the lack of
precipitation for meteorological drought. In SPI, monthly precipitation
is aggregated by a time scale parameter \(k\):
\(x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j),\)
where \(j^\prime\) is the new time index after the aggregation. In this
notation, each spatial location is seaprately aggregated and
precipitation is ummed from \(k\) month back, \(j^\prime - k + 1\), to
the current, \(j^\prime\) period to create the aggregated series,
indexed by \(j^\prime\).

@@

\textbf{Spatial aggregation} {[}TODO{]}

Sometimes

Information from neighbouring spatial locations can also be borrowed
when constructing indexes. This involves interpolating data from
observed locations to the unobserved and aggregating data into different
levels. In the pipeline, spatial aggregation can be written as
\[g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),\] where
\(\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}\) is the spatial
operation parameters and \(d_{\theta}\) is the number of parameter of
\(\theta\). {[}GIVEN AN EXAMPLE{]}

While many indexes are calculated independently on spatial locations

satellite images with different resolutions matching ground and
satellite images data

@@

Distribution fit can be seen as model fitting in its simplest term. It
can be represented by

\[F_{\eta}(x(\mathbf{s}; \mathbf{t})),\] where
\(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\). A
distribution fit typically aims at finding the distribution that best
fit the data. Analysts may start from a pool of candidate distributions
with a chosen fitting method and goodness of fit measure. While it is
useful to find the ultimate best distribution to fits the data, from a
probabilistic perspective, the fitting procedure itself has uncertainty
associated with the data fed and parameter chosen. A reasonable
alternative is to understand how much the index values can vary given
different distributions, fitting methods, and goodness of fit tests, and
whether these variations are negligible in a given application.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi

@@

The construction of an index sometimes needs to consider information
from neighbouring time periods. Temporal processing is a general
operator on the time dimension of the data in the form of
\[f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),\] where
\(\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}\) is the parameters
associated with the temporal operation and \(d_{\psi}\) is the number of
parameter of \(\psi\). An typical example of temporal processing is
aggregation in the drought index SPI to measure the lack of
precipitation for meteorological drought. In SPI, monthly precipitation
is aggregated by a time scale parameter \(k\):
\(x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j),\)
where \(j^\prime\) is the new time index after the aggregation. In this
notation, each spatial location is seaprately aggregated and
precipitation is ummed from \(k\) month back, \(j^\prime - k + 1\), to
the current period, \(j^\prime\), to create the aggregated series,
indexed by \(j^\prime\).

@@

\textbf{Spatial aggregation}

While many indexes calculate independently on spatial locations,
borrowing spatial information can happen at the unobservable locations.
Also, variables can be collected in different formats: some from
satellite imageries as raster data, potentially at different
resolutions, some from ground stations as vector data. These spatial
data need to be merged into a uniform unit to construct indexes. In the
pipeline, spatial aggregation includes operations on the spatial
dimension of the data and can be written as
\[g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t})),\] where
\(\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}\) is the spatial
operation parameters and \(d_{\theta}\) is the number of parameter of
\(\theta\). When one of the variables collected is only available in a
coarser resolution, \(i^\prime\), analysts may choose to reduce other
variables with finer resolution, \(i\), to match the coarser resolution,
as in Figure~\ref{fig-spatial-agg}. This is a spatial aggregation step
and aggregating using the mean can be written as:
\(x(s_{i^\prime};t_{j}) = \bar{x}(s_{i \in i^\prime}; t_j),\), where
\(i \in i^\prime\) includes all the cells from the finer resolution in
the coarser grid. \emph{(Feel like the figure here is more of a visual
illustration of the equation without enough further implication =.=)}

{\centering \includegraphics{paper_files/figure-pdf/fig-spatial-agg-1.pdf}

\caption{\label{fig-spatial-agg}Illustration of aggregating raster data
with finer resolution (right) into coarser resolution (left).}

@@

Distribution fit can be seen as model fitting in its simplest term. It
can be represented by \[F_{\eta}(x(\mathbf{s}; \mathbf{t})),\] where
\(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\). A
distribution fit typically aims at finding the distribution that best
fit the data. Analysts may start from a pool of candidate distributions
with a chosen fitting method and goodness of fit measure. While it is
useful to find the ultimate best distribution to fits the data, from a
probabilistic perspective, the fitting procedure itself has uncertainty
associated with the data fed and parameter chosen. A reasonable
alternative is to understand how much the index values can vary given
different distributions, fitting methods, and goodness of fit tests, and
whether these variations are negligible in a given application.
